---
title: "p8105_hw3_mwd2126"
author: "Michael Denham"
date: "2022-10-17"
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(ggridges)
library(patchwork)
library(p8105.datasets)

```

#Problem 1

First, I begin by loading the data set.

```{r introduce instacart data frame}
data("instacart")
instacart = as_tibble(instacart)
```

The data set `instacart` has `r nrow(instacart)` rows and `r ncol(instacart)` columns. The data set lists items ordered by customers in a long format, with each item getting its own row. Key variables include order_number, aisle_id, department_id, and variables related to the time of the order.

There are `r instacart %>% select(aisle_id) %>% distinct %>% count` aisles. To find which aisles the most items are from, I will build a list of the aisles in order of most items to fewest items.

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

Based on this table, the most items are ordered from the fresh produce aisles. 

Let's look solely at aisles with more than 10,000 items ordered.

```{r}
instacart %>% 
  count(aisle) %>%
  filter(n > 10000) %>% 
  arrange(desc(n)) %>%
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() +
  labs(title = "Number of items ordered in each aisle", x = "Aisle", y = "Number of items") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

Then I'll make a table of the three most popular items in the baking ingredients, dog food care, and packaged vegetables aisles.

```{r}
instacart %>% 
  filter(aisle==c("baking ingredients", "dog food care","packaged vegetables fruits")) %>% 
  group_by(aisle) %>%
  count(product_name) %>%
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>%
  arrange(desc(n)) %>%
  knitr::kable()
```

Next I'll make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week.

```{r}
instacart %>% 
  filter(product_name==c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>% 
  summarize(mean_hr = mean(order_hour_of_day)) %>% 
  spread(order_dow, mean_hr) %>%
  knitr::kable()
```
#Problem 2


